---
title: "ESM206-Lab9-Part2"
author: "Anusha Sridhara"
date: "11/24/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(janitor)
library(broom)
```

## Read in the data
```{r}
lizards <- read_csv(here("data", "lizards.csv"))
```


Get counts of lizards by species: 
```{r}
lizards %>% 
  count(spp) %>% 
  arrange(-n)
```


## Make a subset

- UTST lizards
- Sites of IBPE (enclosure grassland), RABB (rabbit mesquite site), "EAST" (east tarbrush site)
- Tail condition is recorded (W = whole; B=broken)

```{r}
utst <- lizards %>% 
  filter(spp=="UTST") %>% 
  filter(site %in% c("IBPE", "RABB", "EAST")) %>% 
  filter(tail %in% c("B", "W"))
```

## Find counts and proportions of lizards with broken or whole tails at each site

```{r}
#count by site and tail
utst %>% 
  count(site,tail)#is in long format, not pretty in the format of a contingency format
```

```{r}
tail_counts <- utst %>% 
  janitor::tabyl(site, tail)#same info as count but diff format

#proportions by row/ gives percentages
tail_proportions <- tail_counts %>% 
  janitor::adorn_percentages() %>% 
  janitor::adorn_pct_formatting(digits=2) %>% 
  janitor::adorn_ns()
```

Is there a significant association with site and tail breakage
enough evidence to reject the null hypothesis?

## Convert to a contingency table - column to row names 

Use 'column_to_rowname()`

```{r}
lizard_ct <- tail_counts %>% 
  column_to_rownames(var = "site")#now we have sites as row names and not as a variable by itself 
```


## Chi-Square

```{r}
#Are tail breakage and site independent?

lizard_x2 <- chisq.test(lizard_ct)
lizard_x2

lizard_tidy <- broom::tidy(lizard_x2)
```


There is a significant association (i.e, non-independence) between site and lizard tail breakage ($\chi$^2^(`r lizard_tidy$parameter`) = `r lizard_tidy$statistic`, *p* = `r lizard_tidy$p.value`).



















